Categorizing the trajectory dynamics of replay

### 2019-07-19 Article.tex

\documentclass[times, twoside, watermark]{zHenriquesLab-StyleBioRxiv}
\usepackage{blindtext}
\usepackage{amsmath}

@@

To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent states to decompose place cell sequences into categories based on their latent dynamics. Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous. This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories. In addition, the model can function in 2D, avoiding linearization errors on more complicated environments. We demonstrate the utility of this model on simulated and real data of an animal performing a spatial memory task.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=.8\linewidth]{figures/Figure_1}
\caption{Placeholder image of Iris with a long example caption to show justification setting.}
\label{fig:computerNo}
\end{figure}

@@

@@

\section*{Materials and Methods}

>>>>

\documentclass[times, twoside, watermark]{zHenriquesLab-StyleBioRxiv}
\usepackage{blindtext}
\pdfoptionalwaysusepdfpagebox=5
\DeclareMathOperator*{\E}{\mathbb{E}}
\newcommand\norm[1]{\left\lVert#1\right\rVert}

@@

To quantify the structure of replay events, we develop a state space model that uses a combination of discrete and continuous latent state to decompose place cell sequences into categories based on their latent dynamics. Each discrete latent “category” is associated with a type of continuous latent dynamic—hovering in place, spatially fragmented or spatially continuous. This allows for (1) direct comparison between different categories of sequence dynamics, (2) expression of our confidence in one or more categories explaining the data, and (3) characterization of the transitions between categories. In addition, the model can function in 2D, avoiding linearization errors on more complicated environments. We demonstrate the utility of this model on simulated and real data of an animal performing a spatial memory task.

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1.0\linewidth]{figures/Figure2.pdf}
\caption{Simulated data used to fit the encoding model. \textbf{a.} Place fields for each cell are simulated as a Gaussian with a maximum firing rate of 15 spikes / s. \textbf{b.} Position is simulated as a sinusoid to mimic an animal moving back and forth on a linear track. Each cell is simulated as a inhomogeneous Poisson process with a rate that depends on position according to its place field in a.}
\label{fig:computerNo}
\end{figure}

@@

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1.0\linewidth]{figures/Figure3.pdf}
\caption{Model can classify each category of trajectory dynamic (and combinations of dynamics) on simulated data. Each panel shows (from top to bottom) the simulated cell population activity, the probability of the three trajectory dynamics at each time, and the decoded posterior probability of position at each time. \textbf{a.} Continuous. \textbf{b.} Hover. \textbf{c.} Fragmented. \textbf{d.} Hover-Continuous-Hover. \textbf{e.} Fragmented-Continuous-Fragmented. \textbf{f.} Fragmented-Hover-Fragmented.}
\label{fig:computerNo}
\end{figure}

@@

\section*{Materials and Methods}
\subsection*{The Model}
The 
\subsection*{Encoding}
In order to encode each cell's place preference, a generalized linear model was fit to each neuron's spiking activity during movement (defined as time periods when the running speed is greater than 4 cm/s). The model used log link function.

$$\lambda(t) = \exp{(\sum_{i} f_{i}(position)\beta_{i})}$$

A small L2 penalization term $-\lambda\norm{\beta_{i}}_{2}^{2}$ used to prevent model fitting instability when spiking activity is very low. A third degree B-splines basis over position (or tensor products of the B-splines when position is two dimensional) is used to allow firing activity to smoothly vary over position and to reduce the total number of model parameters.
\subsection*{Decoding}

\subsection*{Simulated Data}

\subsection*{Data}

\subsection*{Software and Code availability}
Python code used for analysis and generating figures in the paper is available at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_paper}. Code for the classifier is available in a separate software repository to facilitate code reuse at: \url{https://github.com/Eden-Kramer-Lab/replay_trajectory_classification}. All code is open-source and licensed under the MIT Software License.
